---
title: "Intro to Monte-Carlo"
output: html_notebook
---

This R Notebook is intended to demonstrate various concepts from probability and Monte-Carlo simulation in R. It is intended for use in the "Financial Analytics" course in the MSc in Business Analytics program at Imperial College Business School.


### Simulating from the Uniform Distribution
First let's simulate from the U(a,b) distribution.
```{r}
n=10000
a = 2
b = 5
U = runif(n, a, b)   # Don't need to specify min and max arguments if simulating from U(0,1)
U[1:10]
```
### Setting the Seed
It's often a good idea (for reproducibility reasons!) to set the seed of the random number generator.
```{r}
set.seed(42); 
u1 = runif(10)
u1
```
```{r}
set.seed(42) 
u2 = runif(10) 
u2 # should be identical to u1 because we used identical seeds for each generation 
```

```{r}
u3 = runif(10) 
u3 # But this will be different
```
This works for all random numbers
```{r}
set.seed(42) 
n1 = rnorm(5)
n1

set.seed(42) 
n2 = rnorm(5) 
n2 # should be identical to n1 because we used identical seeds for each generation 

set.seed(8765)
n3 = rnorm(5) 
n3 # But this will be different
```

### An Aside: Plotting CDFs and PMFs / PDFs
Our first curve plots the empirical CDF. The empirical CDF is the CDF that places a probability weight of 1/n on each of the n simulated data-points. We will also add the true uniform CDF to the plot.
```{r}
n=100
U = runif(n, a, b)
plot(ecdf(U))
curve(punif(x, a, b), add=TRUE, lty=2,col=3)   # The uniform CDF
```
We can also plot the fitted density of the data using the **density** function as well as PDF of the true underlying distribution using the **curve** function. Note the fitted density doesn't know the true PDF and so it may place some probability outside the [a,b] interval which is actually where all the probability resides.
```{r}
plot(density(U)) 
curve(dunif(x, a, b), add=TRUE, col=3)  
```


### Simulating from the Normal Distribution
```{r}
mu = 3
sigma = 2
n = 100000
NormData = rnorm(n, mean=mu, sd=sigma);
plot(density(NormData)) # the PDF estimated from the simulated data
curve(dnorm(x, mean=mu, sd=sigma), add=TRUE, col=3)   # add the true normal PDF curve
```
May also want to see a histogram of the simulated data
```{r}
hist(NormData, col = "wheat", label = T)
```

### Simulating from a Discrete Distribution
Want to simulate a value of X which takes on the value X[i] with probability Probs[i]
```{r}
Probs = c(0.35,.15,.4,.1)
X = c(1,2,3,4);   
P_cum = cumsum(Probs) # Let's compute the cumulative probability vector
P_cum
```
The following code now simulates a point from this discrete distribution. Depending on the distribution there may be (much more) efficient ways of doing this.
```{r}
counter = 1;
r = runif(1);
while(r > P_cum[counter]){   
  counter = counter + 1;
}
X[counter]
```

### Generating $\mbox{MN}_m(\mu, \Sigma)$ Random Vectors
```{r}
mu = c(0,0)
rho = .5
Sigma = matrix(c(1,rho,rho,1),nrow=2)
Sigma

NumSamples = 10000
library(MASS)
NormData = mvrnorm(NumSamples, mu, Sigma)
cor(NormData)  # checking that the sample correlation matrix agrees with the true correlation matrix
```
Note that in practice you will generally want to **compute confidence intervals** for whatever quantity you're estimating.