---
title: "Cheetsheet_2"
author: "Qian Zhang"
date: "2020年12月10日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

### ggplot: Continuous univariate
```{r}
ggplot(data = mpg, mapping = aes(x = hwy)) +
geom_histogram()
ggplot(data = mpg, mapping = aes(x = hwy)) +
geom_freqpoly()
ggplot(data = mpg, mapping = aes(x = hwy)) +
geom_density()
```

### ggplot: density and area
```{r}
ggplot(data = mpg, aes(x = hwy)) +
geom_density(alpha = .2, fill= "#00BFC4", color = 0,adjust = 1) +
geom_line(stat='density') 
```

### ggplot: comparing distribution
```{r}
mpg %>%
filter(manufacturer %in% c("dodge", "toyota", "volkswagen")) %>%
ggplot(aes(x = hwy, color = manufacturer)) +
geom_density(adjust = 0.5)
```

## ggplot: categorical univariate
```{r}
ggplot(data = mpg, mapping = aes(x = class)) +
geom_bar()
```

## ggplot: categorical univariate with forcats funcions
fct_reorder() reordering a factor by another variable
fct_infreq() reodering a factor by the frequency of value
fct_relevel() changing the order of a factor by hand
fct_lump() collapsing the least/most frequent values of a facotr into “other”
fct_rev() reverse order of factor levels
```{r}
ggplot(data = mpg, mapping = aes(x = fct_infreq(class))) +
geom_bar()
```

## ggplot: scatter plot for continuous bivariate
```{r}
ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = hwy))
```

## ggplot: scatter plot for continuous bivariate
```{r}
ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = hwy))
```

## ggplot: smooth line for continuous bivariate 
```{r}
ggplot(data = mpg) +
geom_smooth(mapping = aes(x = displ, y = hwy))
```

## ggplot: box plot for continuous bivariate 
```{r}
ggplot(data = mpg, aes(x = class, y = hwy)) +
geom_boxplot()
```

## ggplot: violin plot for continuous bivariate 
```{r}
ggplot(data = mpg, aes(x = class, y = hwy)) +
geom_violin()
```

## ggplot: scatter plot and violin plot for continuous bivariate 
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
geom_point() +
geom_smooth()
```

## ggplot: Grid of plots
```{r}
library(gridExtra)
p1 <- ggplot(data = mpg) + geom_bar(aes(x = hwy))
p2 <- ggplot(data = mpg) + geom_point(aes(x = hwy, y = displ), alpha = 0.5)
grid.arrange(p1, p2, ncol = 2)
```

## ggplot: patchwork package
```{r}
library(patchwork)
p1 <- ggplot(data = mpg) + geom_bar(aes(x = hwy))
p2 <- ggplot(data = mpg) + geom_point(aes(x = hwy, y = displ), alpha = 0.5)
p1 + p2
```

## density plot for different classes
```{r}
mpg %>%
ggplot(aes(x = hwy, color = class)) +
geom_density(aes(y = ..count..)) +
facet_wrap(vars(class), nrow = 2)
```

## four options for geom_bar
1. stack : default op5on to create a stacked barcharts
2. identity : un-stacked, draws each object exactly where it falls in the context of the graph. This op5on is not very helpful as bars overlap
3. dodge : avoids overlapping bars by placing beside one another
4. fill : works like stack but visualise the propor5ons across groups
```{r}
library(patchwork)
p1 <- ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = clarity), position = "stack")
p2 <- ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = clarity), position = "identity")
p3 <-  ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
p4 <- ggplot(data = diamonds) +  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
p1 + p2 + p3 + p4
 
```

## avoiding overploeng by adding small amount of random noise to each point to spread the overlapping points out e.g. scatter plot
```{r}
library(patchwork)
p1 <- ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy), position = "jitter", alpha = 0.5)
p2 <- ggplot(data = mpg) + geom_jitter(aes(x = displ, y = hwy), alpha = 0.5)
p1 + p2
```

## Practice 1
```{r}
ggplot(mpg,aes(displ,hwy))+
  geom_point()+
  geom_smooth(method='loess',formula = 'y~x')
```

## Practice 2
```{r}
ggplot(mpg,aes(displ,hwy))+
  geom_point()+
  geom_smooth(aes(group=drv),method='loess',formula = 'y~x',se=F)
```

## Practice 3
```{r}
ggplot(mpg,aes(displ,hwy))+
  geom_point(aes(color=drv))+
  geom_smooth(aes(group=drv,color=drv),method='loess',formula = 'y~x',se=F)
```

## Practice 4
```{r}
ggplot(mpg,aes(displ,hwy))+
  geom_point(colour="white",size=4)+
  geom_point(aes(colour=drv))
```

## Practice 5
```{r}
ggplot(mpg,aes(x=displ))+
  geom_density(aes(y=..count..,fill=drv),alpha=0.2,color=0)+
  geom_line(aes(y=..count..,group=drv),stat="density")
```

## Practice 6
```{r}
ggplot(mpg,aes(class,hwy))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(aes(group=class),width=0.25,color="blue")
```

## Practice 7
```{r}
p1 <- ggplot(mpg) +
  geom_point(aes(x = hwy, y = displ), alpha = 0.5) 
p2 <- ggplot(mpg) +
  geom_histogram(aes(x = hwy), binwidth = 2, na.rm = T) 

p3 <- ggplot(mpg) +
  geom_histogram(aes(x = displ), binwidth = 0.5, na.rm = T) 

p1+p2+p3
```

