---
title: "In-class Exercise Lecture 8"
author: "Your CID number here!"
date: "17/11/2020"
output: html_document
---

```{r setup, include=FALSE}
# Load the tidyverse packages
library(tidyverse)
library(nycflights13)
```

## Instruction

* Complete this RMarkdown file and "Knit" the HTML output (saved as **Lecture_8_YOUR_CID_NUMBER.html**) and upload to the Hub before the deadline.
* Change the author property in the YAML header above to your CID number. 

* For this exercise, we will use `flights` dataset from `nycflights13` package. Install the `nycflights13` if you have not installed yet. This is a dataset of all flights that departed NYC (JFK, LGA or EWR) in 2013. 
* To complete this exercise, make sure you have installed the tidyverse package. `install.packages("tidyverse")`

### Filtering

Find the number of flights that:
```{r eval = F}
# Had an arrival delay of two or more hours
# [1] 10200

# Flew to Houston (IAH or HOU)
# [1] 9313

# Were operated by United, American, or Delta
# [1] 139504

# Departed in summer (July, August, and September) and have missing departure time values
# [1] 1878

# Arrived more than two hours late, but didn’t leave late
# [1] 29

# Were delayed by at least an hour, but made up over 30 minutes in flight
# [1] 2074
```
## Sorting
List the first 3 rows of the resulting dataframe. The outputs shown here only list the first 9 columns and may not show the column that was sorted on. 

```{r eval=F}
# Sort flights to find the most delayed flights

#    year month   day dep_time sched_dep_time dep_delay arr_time sched_arr_time arr_delay
#   <int> <int> <int>    <int>          <int>     <dbl>    <int>          <int>     <dbl>
# 1  2013     1     9      641            900      1301     1242           1530      1272
# 2  2013     6    15     1432           1935      1137     1607           2120      1127
# 3  2013     1    10     1121           1635      1126     1239           1810      1109


# Find the flights that left earliest

#    year month   day dep_time sched_dep_time dep_delay arr_time sched_arr_time arr_delay
#   <int> <int> <int>    <int>          <int>     <dbl>    <int>          <int>     <dbl>
# 1  2013     1    13        1           2249        72      108           2357        71
# 2  2013     1    31        1           2100       181      124           2225       179
# 3  2013    11    13        1           2359         2      442            440         2


# Which flights traveled the longest in terms of distance?

#    year month   day dep_time sched_dep_time dep_delay arr_time sched_arr_time arr_delay carrier flight
#   <int> <int> <int>    <int>          <int>     <dbl>    <int>          <int>     <dbl> <chr>    <int>
# 1  2013     1     1      857            900        -3     1516           1530       -14 HA          51
# 2  2013     1     2      909            900         9     1525           1530        -5 HA          51
# 3  2013     1     3      914            900        14     1504           1530       -26 HA          51

# Which traveled the shortest in terms of distance?

#    year month   day dep_time sched_dep_time dep_delay arr_time sched_arr_time arr_delay carrier flight
#   <int> <int> <int>    <int>          <int>     <dbl>    <int>          <int>     <dbl> <chr>    <int>
# 1  2013     7    27       NA            106        NA       NA            245        NA US        1632
# 2  2013     1     3     2127           2129        -2     2222           2224        -2 EV        3833
# 3  2013     1     4     1240           1200        40     1333           1306        27 EV        4193


```

## Grouped summaries

```{r eval = F}

# Which carrier had the largest mean departure delay?  Smallest?
# F9       20.2
# US       3.78


# Which carrier had the largest difference between their max and min departure delay?
#   carrier   max   min delta
#   <chr>   <dbl> <dbl> <dbl>
# 1 HA       1301   -16  1317


# Which month has the largest variance for arrival delays? (use standard deviation)
# July


```

## Pipe

Using the pipe operator follow these steps with the flight data:

* filter out missing values that exist in the tail number and arrival delay variables
* group the data by tail number
* calculate the average arrival delay and the number of observations by tail number
* plot average arrival delay vs. number of observations

```{r}
# Your code

```


Using the pipe operator follow these steps with the flight data:

* group by month and carrier
* compute the average arrival delay
* plot a line chart with month on the x-axis, average delay on the y-axis
* colour-code each line by carrier or try faceting by carrier

```{r}
# Your code

```

