---
title: "In-class Exercise Lecture 10"
author: "Your CID number here!"
date: "26/11/2020"
output: html_document
---

```{r setup, include=FALSE}
# Load the tidyverse packages
library(tidyverse)
library(plotly)
```

## Instruction

* Complete this RMarkdown file and "Knit" the HTML output (saved as **Lecture10_YOUR_CID_NUMBER.html**) and upload to the Hub before the deadline.
* Change the author property in the YAML header above to your CID number. 
* Your task is to create static/interactive charts based on 3 different datasets, and refine the charts.


## Plot 1
Examine the `ggplot2::economics` dataset first, and come up with a hypothesis, or an idea for which attributes or patterns to compare. Then, go to [from Data to Viz website](https://www.data-to-viz.com) for an inspiration and recreate a chart using this dataset. Document your thought process below in Markdown.

```{r}
# your code here

```

#### Hypothesis
*

#### Insights
*
*
*

## Plot 2
Examine the `ggplot2::diamonds` dataset first, and come up with a hypothesis, or an idea for which attributes or patterns to compare. Then, create a static `ggplot2` chart of your choice. Then, apply `plotly::ggplotly()` function to convert ggplot object to plotly graph. Document your thought process below in Markdown

```{r}
# your code here

```

#### Hypothesis
*

#### Insights
*
*
*

## Plot 3 
Examine the `ggplot2::txhousing` dataset first, and come up with a hypothesis, or an idea for which attributes or patterns to compare. Then, go to [plotly website](https://plotly.com/r/) for an inspiration and recreate an **interactive** chart using this dataset. Document your thought process below in Markdown.


```{r}
# your code here


```

#### Hypothesis
*

#### Insights
*
*
*
