---
title: "Assignment 2: SCANPRO Model"
output: html_document
---


```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(dplyr)
library(readxl)
library(kableExtra)

```

<br/>
<br/>


Now it is time to apply what you’ve learned to some exciting, real-world data!

In file scanpro_beauty.xlsx, you will be able to get access to monthly sales data of a world leading beauty product in a regional market in Asia for a 51-month period. 

In the dataset you will find the following variables:

* **month**: monthly index (1 to 51)

* **sales**: regional monthly sales volume (in thousands) of the beauty product within a specific month.

* **price**: monthly average price (in cents) of the beauty product across all retailers in the region.

* **presence**: monthly retailer shelf presence index (i.e., a measure of shelf visibility developed by the brand internally; the higher the value, the higher the shelf visibility).

* **media**: monthly media coverage index (i.e., a comprehensive measure of overall intensity of media exposure across all marketing channels of the brand; the higher the value, the more intensive the media exposure)

<br/>

Here is a pre-view of part of the dataset:

```{r echo=FALSE, message=FALSE, warning=FALSE}

setwd("~/Downloads/scanpro")

scanpro_beauty <- read_excel("scanpro_beauty.xlsx")

kable(head(scanpro_beauty[,1:5]),caption = "Table 1 Sample Monthly Sales Data")%>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
  
```

<br/>

### Task:

<br/>

* Assess the effectiveness of various types of marketing efforts in driving up monthly sales by fitting a SCAN*PRO model.

* Interpret model parameters with marketing insights.

* Comment on possible ways to improve model performance.
  
<br/>
<br/>


