---
title: "R Notebook"
output: html_notebook
---

The olivettifaces.mat (Matlab) file contains a 4096 x 400 matrix called "faces" that represents 400 greyscale faces (10 images for each of 40 people). See information on this dataset at http://scikit-learn.org/stable/datasets/olivetti_faces.html
In particular : "There are ten different images of each of 40 distinct subjects. For some subjects, the images were taken at different times, varying the lighting, facial expressions (open / closed eyes, smiling / not smiling) and facial details (glasses / no glasses). All the images were taken against a dark homogeneous background with the subjects in an upright, frontal position (with tolerance for some side movement)." The .mat file also contains variables p, u nd v but they can be ignored.
```{r}
library(R.matlab)
data = readMat("olivettifaces.mat")  
faces = data$faces
dim(faces)
NumFaces = length(faces[1,])
NumPixels = length(faces[,1]) 
```
```{r}
NumFacesPCA = 400;  # Number of images we will use for PCA
set.seed(1)
index = sample(1:NumFaces, NumFacesPCA);
TrainImages = faces[,index]
```


```{r}
library(plotrix)
n = 2
par(mfrow = c(n,n),     # 2x2 layout
    oma = c(0, 0, 0, 0), # controls rows for text at outer  margins
    mar = c(1, 1, 0, 0), # space for one row of text at ticks and to separate plots
    mgp = c(2, 1, 0),    # axis label at 2 rows distance, tick labels at 1 row
    xpd = NA)  
for (i in 1:n^2){
  Face = matrix(TrainImages[,i],sqrt(NumPixels),byrow=FALSE)
  color2D.matplot(Face,axes=FALSE)
}
```
Let's compute the mean face and plot it. 
```{r}
mean_image = apply(TrainImages, 1, mean)
mean_Face = matrix(mean_image,sqrt(NumPixels),byrow=FALSE)
color2D.matplot(mean_Face,axes=FALSE)
```

