---
title: "Cheetsheet_4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(viridis)
library(ggplot2)
library(RColorBrewer)
library(dplyr)
```

## Fill & colour scale - Discrete
scale_*_discrete(): Colours evenly spaced around the colour wheel (same as hue)
scale_*_hue(): Colours evenly spread around the colour wheel
scale_*_grey(): Greyscale palette
scale_*_viridis_d(): Viridis palettes
scale_*_brewer(): ColorBrewer palettes
scale_*_manual(): Manually specified colours

## Fill & colour scale - Continuous
scale_*_gradient(): Two-colour gradient
scale_*_gradient2(): Gradient with a middle colour and tow colours that diverge from it
scale_*_gradientn(): Gradient with n colour, equally spaced
scale_*_viridis_c(): Viridis pale:es



# Exploratory data analysis: Organ data

## Organ data: 
```{r}
organdata <- socviz::organdata
ggplot(data = organdata, mapping = aes(x = year, y = donors)) +
 geom_point()
```

## Organ data: remove empty data
```{r}
ggplot(data = organdata, mapping = aes(x = year, y = donors)) +
 geom_point(na.rm = T)
```

## Organ data: draw lines by every group of countries
```{r}
ggplot(data = organdata, mapping = aes(x = year, y = donors)) +
 geom_point(na.rm = T) +
 geom_line(aes(group = country), na.rm = T)
```

## Organ data: draw lines by every group of countries in different plots
```{r}
ggplot(data = organdata, mapping = aes(x = year, y = donors)) +
  geom_point(na.rm = T, size = 0.5) +
  geom_line(aes(group = country), na.rm = T) +
  facet_wrap(~ country)
```

## Organ data: boxplot
```{r}
ggplot(data = organdata, mapping = aes(x = country, y = donors)) +
  geom_boxplot(na.rm = T)
```

## Organ data: boxplot flipped
```{r}
ggplot(data = organdata, mapping = aes(x = country, y = donors)) +
  geom_boxplot(na.rm = T) +
  coord_flip()
```

## Organ data: boxplot ordered
```{r}
ggplot(data = organdata,
       mapping = aes(x = reorder(country, donors, na.rm = T), y = donors)) +
  geom_boxplot(na.rm = T) +
  coord_flip()
```

## Organ data: Data transforma8on
```{r}
by_country <- organdata %>%
  group_by(consent_law, country) %>%
  summarize(donors_mean= mean(donors, na.rm = TRUE),
            donors_sd = sd(donors, na.rm = TRUE),
            gdp_mean = mean(gdp, na.rm = TRUE),
            health_mean = mean(health, na.rm = TRUE),
            roads_mean = mean(roads, na.rm = TRUE),
            cerebvas_mean = mean(cerebvas, na.rm = TRUE))
```

## Organ data: Cleveland dotplot
```{r}
ggplot(data = by_country,
       mapping = aes(x = donors_mean, y = reorder(country, donors_mean),
                     color = consent_law)) +
  geom_point(size=3) +
  labs(x = "Donor Procurement Rate", y = "", color = "Consent Law") +
  theme(legend.position="top")
```

## Organ data: plot texts
```{r}
library(ggrepel)
ggplot(data = by_country,mapping = aes(x = roads_mean, y = donors_mean)) +
  geom_point() +
  geom_text_repel(mapping = aes(label = country))
```

## Organ data: label outliers
```{r}
ggplot(data = by_country,mapping = aes(x = roads_mean, y = donors_mean)) +
  geom_point() +
  geom_text_repel(data = subset(by_country, roads_mean > 140 | donors_mean < 15),
                  mapping = aes(label = country))
```

# Exploratory data analysis: Movie data
```{r}
movies<-ggplot2movies::movies
```

## Movie Data: Assessing the distribution
```{r}
ggplot(data = movies, mapping = aes(x = length))+
  geom_histogram()
```

## Movie Data: limit the x axis
```{r}
ggplot(data = movies, mapping = aes(x = length))+
  geom_histogram(binwidth = 5) +
  coord_cartesian(xlim = c(0, 60*4))
```

## Movie Data: Defining “long” and find how many long movies
```{r}
movies %>%
  count(cut_width(length, 5)) %>%
  mutate(cum_pct = cumsum(n)/sum(n)) %>%
  filter(cum_pct > .95) %>%
  summarise(sum(n))
```

## Movie Data: Top 5 longest 
```{r}
movies %>%
  arrange(desc(length)) %>%
  top_n(5, wt = length)
```

## Movie Data: Labelling long movies
```{r}
movies %>%
  select(1:3) %>%
  mutate(Long = length >= 122.5)
```

# Exploratory data analysis: Short films

## Movie Data: Defining short movies
```{r}
ggplot(data = movies, mapping = aes(length)) +
  geom_histogram(binwidth = 2) +
  coord_cartesian(xlim = c(0, 122.5),
                  ylim = c(0, 3000))
```

## Movie Data: how many short films?
```{r}
movies %>%
  filter(length <= 40) %>%
  summarise(n())
```

